\section{Phishing \& Cyberwar}

\pgra{Definition:} Phishing is the criminally fraudulent process of attempting to acquire sensitive information such as usernames, passwords, and credit card details by masquerading as a trustworthy entity in an electronic communication.

\pgra{Phishing -- PayPal Summary:} Original spam mail: sent through open relay mailserver in UK. Web Site: refers to online-security-paypal.com, points to French web forwarding service, page really resides on hacked Web server in UK operated by Spanish (real estate) customer. Domainname: Domain owner in NY, registered at day of the phishing mail, using a French registrar.

\pgra{Mule Recruitment -- money laundering:} Mails sent out in the form of an attractive job offer allowing the user to work at home. In reality, he launders money for cybercriminals by
\bitem{
\item[...] receiving money transfers to personal account,
\item[...] keeping 10\% from each transaction,
\item[...] forwarding the remainder through Money Gram or Western Union
}

The company behind the scam ``Elx Trading'' claims to do background checks. It is proclaimed to be in St. Petersburg, Russia, its web site is hosted in Hong Kong and later in the U.S. The domain registration is in Palestinian territories, whereas the phone number was for Iraq, etc. $\to$ The tricked user ends up with the liability for having laundered money. The criminals themselves are hard to track and shut down permanently.

\pgra{Rock phish kit:} Host multiple fake sites for different brands on a single host. April 2007: 97.3\% new Phishing hosts appear to be kit-based.

\pgra{Conclusions:}
\bitem{
\item Phishing is an organized crime, the reception of a phishing email is the first phase of an international money-making machine
\item Cybercriminals are highly professional and organized, their business model works
\item Poor system security allows cyber criminals to start and propagate their business
\item Any technical, social, psychological, system, or process weakness is going to be exploited in an efficient and (potentially) automated way
}

\pgra{Social Networks}
Mass communication, event coordination. Distribution of cyber arms and coordinated attacks are possible (cyber-jihad). Mass protest. Mail bombers, network flooders. Opt-in Botnet.\\
\textit{￼Attack ramifications:} Economic exhaustion, public disinformation, flooding of non-internet system \\
\textit{Defense strategies:} Social Network Sites bear great burden of protecting organizations from being targeted by their own members. Monitoring, content sanitization (censorship), user agreements, etc.

\pgra{Cyber warfare:} Internet was built for convenience, reliability, not for security. Cyberspace has become the 5th domain of warfare. Everyone has access to the latest weapon system technology. Favors early, pre-emptive attacks. Plausible deniability. \\ \textit{Aliter:} It is currently under discussion which actions constitue cyber warfare and to which extend cyber warfare can result in war crimes.

\textit{The historic monopoly of states to access the latest weapon technology is broken.}

\section{General Information}

\pgra{The OSI and TCP/IP layers:} layer conversion:

\begin{tabular}{|p{0.45\linewidth}|p{0.45\linewidth}|}
\hline
\textbf{OSI} & \textbf{TCP/IP} \\
\hline
\hline
Data link layer & Link layer \\
\hline
Network layer & IP layer \\
\hline
Transport layer & TCP, UDP layer \\
\hline
Session layer & \multirow{3}{*}{Application layer} \\
Presentation layer & \\
Application layer & \\
\hline
\end{tabular}
